Perturbation Theory in the Complex Plane: Exceptional Points and Where to Find Them

### 2020-12-02 Manuscript/EPAWTFT.tex
saving work in Shanks

\titou{T2: $N$ is defined as the number of electrons. Maybe we should use $n$ instead?}
Consider the partial sums
$S_N = \sum_{k=0}^{N} a_k$
defined from the truncated summation of an infinite series 
$    S = \sum_{k=0}^{\infty} a_k$.
If the series converges, then the partial sums will tend to the exact result 
\begin{equation}
\lim_{N\rightarrow \infty} S_N = S. 
\end{equation}
The Shanks transformation attempts to generate increasingly accurate estimates of this
limit by defining a new series as
\begin{equation}
    T(S_N) = \frac{S_{N+1} S_{N-1} - S_{N}^2}{S_{N+1} + S_{N-1} - 2 S_{N}}.
\end{equation}
This series can converge faster than the original partial sums and can thus provide greater
accuracy using only the first few terms in the series.
However, it is designed to accelerate exponentially converging partial sums with 
the approximate form
\begin{equation}
    S_N \approx S + a\,b^N.
\end{equation}
Furthermore, while this transformation can accelerate the convergence of a series, 
there is no guarantee that this acceleration will be fast enough to significantly
improve the accuracy of low-order approximations.

\hugh{To the best of our knowledge, the Shanks transformation has never previously been applied
to the acceleration of the MP series.
We have therefore applied it to the convergent Taylor series, Pad\'e approximants, and quadratic
approximants for RMP and UMP in the symmetric Hubbard dimer. 
The UMP approximants converge too slowly for the Shanks transformation
to provide any improvement, even in the case where the quadratic approximants are already 
very accurate.
In contrast, acceleration of the diagonal Pad\'e approximants for the RMP cases 
can significantly improve the estimate of the energy using low-order perturbation terms, 
as shown in Table~\ref{tab:RMP_shank}.
Even though the RMP series diverges at $U/t = 4.5$, the combination
of diagonal Pad\'e approximants with the Shanks transformation reduces the absolute error of 
the best energy estimate to 0.002\,\%.}

\begin{table}[th]
	\caption{
    \hugh{%
    Acceleration of the diagonal Pad\'e approximant sequence for the RMP energy
    using the Shanks transformation.
    }
	\label{tab:RMP_shank}}
	\begin{ruledtabular}
		\begin{tabular}{lcccc}
            &			&	&	\mc{2}{c}{$E_{-}(\lambda = 1)$} \\
												\cline{4-5} 
            Method  &   Degree	&  Series Term &   $U/t = 3.5$ & $U/t = 4.5$ \\
			\hline
            Pad\'e	&	[1/1]	& $S_1$        &  $-1.61111$   &  $-2.64286$ \\
                    &   [2/2]   & $S_2$        &  $-0.82124$   &  $-0.48446$ \\
                    &   [3/3]   & $S_3$        &  $-0.91995$   &  $-0.81929$ \\
                    &   [4/4]   & $S_4$        &  $-0.90579$   &  $-0.74866$ \\
                    &   [5/5]   & $S_5$        &  $-0.90778$   &  $-0.76277$ \\
            \hline
            Shanks  &	        &  $T(S_2)$    &  $-0.90898$   &  $-0.77432$ \\
                    &           &  $T(S_3)$    &  $-0.90757$   &  $-0.76096$ \\
                    &           &  $T(S_4)$    &  $-0.90753$   &  $-0.76042$ \\
			\hline
            Exact	&	        & 	           &  $-0.90754$	&  $-0.76040$ \\
		\end{tabular}
	\end{ruledtabular}
\end{table}

>>>>

Consider the partial sums
$S_n = \sum_{k=0}^{n} s_k$
defined from the truncated summation of an infinite series 
$S = \sum_{k=0}^{\infty} s_k$.
If the series converges, then the partial sums will tend to the exact result 
\begin{equation}
	\lim_{n \to \infty} S_n = S. 
\end{equation}
The Shanks transformation attempts to generate increasingly accurate estimates of this
limit by defining a new series as
\begin{equation}
    T(S_n) = \frac{S_{n+1} S_{n-1} - S_{n}^2}{S_{n+1} + S_{n-1} - 2 S_{n}}.
\end{equation}
This series can converge faster than the original partial sums and can thus provide greater
accuracy using only the first few terms in the series.
However, it is designed to accelerate \titou{exponentially?} converging partial sums with 
the approximate form
\begin{equation}
    S_n \approx S + \alpha\,\beta^n.
\end{equation}
Furthermore, while this transformation can accelerate the convergence of a series, 
there is no guarantee that this acceleration will be fast enough to significantly
improve the accuracy of low-order approximations.

To the best of our knowledge, the Shanks transformation has never previously been applied
to the acceleration of the MP series.
We have therefore applied it to the convergent Taylor series, Pad\'e approximants, and quadratic
approximants for RMP and UMP in the symmetric Hubbard dimer. 
The UMP approximants converge too slowly for the Shanks transformation
to provide any improvement, even in the case where the quadratic approximants are already 
very accurate.
In contrast, acceleration of the diagonal Pad\'e approximants for the RMP cases 
can significantly improve the estimate of the energy using low-order perturbation terms, 
as shown in Table~\ref{tab:RMP_shank}.
Even though the RMP series diverges at $U/t = 4.5$, the combination
of diagonal Pad\'e approximants with the Shanks transformation reduces the absolute error of 
the best energy estimate to 0.002\,\%.

\begin{table}[th]
	\caption{
    Acceleration of the diagonal Pad\'e approximant sequence for the RMP energy
    using the Shanks transformation.
	\label{tab:RMP_shank}}
	\begin{ruledtabular}
		\begin{tabular}{lcccc}
            &			&	&	\mc{2}{c}{$E_{-}(\lambda = 1)$} \\
												\cline{4-5} 
            Method  &   Degree	&  Series Term &   $U/t = 3.5$ & $U/t = 4.5$ \\
			\hline
            Pad\'e	&	[1/1]	& $S_1$        &  $-1.61111$   &  $-2.64286$ \\
                    &   [2/2]   & $S_2$        &  $-0.82124$   &  $-0.48446$ \\
                    &   [3/3]   & $S_3$        &  $-0.91995$   &  $-0.81929$ \\
                    &   [4/4]   & $S_4$        &  $-0.90579$   &  $-0.74866$ \\
                    &   [5/5]   & $S_5$        &  $-0.90778$   &  $-0.76277$ \\
            \hline
            Shanks  &	        &  $T(S_2)$    &  $-0.90898$   &  $-0.77432$ \\
                    &           &  $T(S_3)$    &  $-0.90757$   &  $-0.76096$ \\
                    &           &  $T(S_4)$    &  $-0.90753$   &  $-0.76042$ \\
			\hline
            Exact	&	        & 	           &  $-0.90754$	&  $-0.76040$ \\
		\end{tabular}
	\end{ruledtabular}
\end{table}
